Density-dependent selection with strictly relative contests

### 2017-11-16 manuscript.tex
l

There are a variety of different measures of fitness. Some widely used examples in evolutionary ecology are expected lifetime reproductive ratio $R_0$, intrinsic growth rate $r$, saturation population density (often labeled ``$K$'') \citep{benton_2000}, and invasion fitness \citep{metz_1992}. In addition, ``relative fitness'' is the standard in much of evolutionary biology, particularly evolutionary genetics, where attention is generally restricted to relative genotypic proportions \cite[pp. 468]{barton_2007}. This variety is not problematic by itself because different fitness measures serve different purposes. But ultimately any measure of fitness should be grounded in the processes of birth and death which govern population ecology and demography \citep{metcalf_2007}. While this connection is apparent for absolute fitness measures like $r$ and $K$, relative fitness seems largely divorced from population ecology.

@@

As discussed in the Introduction, MacArthur's general argument starts from Eq.~\eqref{eq:macgeneral} and then hinges on the shape of the nullclines $f_1(n_1,n_2)=0$ and $f_2(n_1,n_2)=0$. The symbol ``$K$'' enters the argument as the general label for the four intersection points of the nullclines with the $n_1$ and $n_2$ axes, specifically $f_1(K_{11},0)=0$, $f_1(0,K_{12})=0$, $f_2(0,K_{22})=0$ and $f_2(K_{21},0)=0$. These $K$ values determine whether a region of higher-density growth exists for each type, provided that the nullclines are close to being straight lines.

>>>>

There are a variety of different measures of fitness. Some widely used examples in evolutionary ecology are expected lifetime reproductive ratio $R_0$, intrinsic growth rate $r$, saturation population density (often labeled ``$K$'') \citep{benton_2000}, and invasion fitness \citep{metz_1992}. In addition, ``relative fitness'' is the standard in much of evolutionary biology, particularly evolutionary genetics, where attention is generally restricted to relative genotypic proportions \cite[pp. 468]{barton_2007}. This variety is not necessarily problematic in itself, because different fitness may be more useful in different circumstances. But any measure of fitness should ultimately be grounded in the processes of birth and death which govern population ecology and demography \citep{metcalf_2007,doebeli_2017}. While this grounding is apparent for absolute fitness measures like $r$ and $K$, relative fitness seems largely divorced from population ecology.

@@

In the Introduction we described MacArthur's general conclusions about the behaviour of Eq.~\eqref{eq:macgeneral} based on the shape of the nullclines $f_1(n_1,n_2)=0$ and $f_2(n_1,n_2)=0$, but did not specify the role of ``$K$''.

The symbol ``$K$'' enters the argument as the general label for the four intersection points of the nullclines with the $n_1$ and $n_2$ axes, specifically $f_1(K_{11},0)=0$, $f_1(0,K_{12})=0$, $f_2(0,K_{22})=0$ and $f_2(K_{21},0)=0$. These $K$ values determine whether a region of higher-density growth exists for each type, provided that the nullclines are close to being straight lines.
